Fast Sorting on Modern Computers

### 2005-08-31 docs/tech_report/report.tex
did the abstract. removed comment from the latex files

The other main work described in this report is an analysis of the
behaviour of sorting algorithms on modern caches. Over the last decade
there has been considerable interest in optimizing sorting algorithms
to reduce the number of cache misses. We experimentally study the
cache performance of both classical sorting algorithms, and a variety
of cache-optimized algorithms, primarily those proposed by LaMarca and
Ladner. Our experiments cover a much wider range of algorithms than
other work, including the $O(N^2)$ sorts and shellsort, all within a
single framework. We discover a number of new results, particularly
relating to the branch prediction behaviour of cache-optimized sorts.

We also developed a number of other improvements to the algorithms, such as
removing the need for a sentinel in heapsort. Overall, we found that a
cache-optimized radixsort was the fastest algorithm for all large arrays; the
absence of comparison branches means that the algorithm causes almost no branch
mispredictions.

% TODO add copy back
\end{abstract}
\end{spacedpars}

@@

%TODO \chapter{Software Predictor Result Listing}
%\input software_predictor.tex

>>>>

The other main work described in this report is an analysis of the behaviour of
sorting algorithms on modern caches. Over the last decade there has been
considerable interest in optimizing sorting algorithms to reduce the number of
cache misses. We experimentally study the cache performance of both classical
sorting algorithms, and a variety of cache-optimized algorithms proposed by
LaMarca and Ladner. Our experiments cover a much wider range of algorithms than
other work, including the $O(N^2)$ sorts, radixsort and shellsort, all within a
single framework. We discover a number of new results, particularly relating to
the branch prediction behaviour of cache-optimized sorts.

We also developed a number of other improvements to the algorithms, such as
removing the need for a sentinel in classical heapsort. Overall, we found that a
cache-optimized radixsort was the fastest sort in our study; the absence of
comparison branches means that the algorithm causes almost no branch
mispredictions.

\end{abstract}
\end{spacedpars}

@@

%\input software_predictor.tex
